---
title: "pvcurve_model"
author: "me"
date: "`r Sys.Date()`"
output: html_document
bibliography: autopv.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fitting PV parameters not by eye(AutoPV)

Pressure volume curves are created by repeated measurements of leaf fresh mass and leaf water potential( $\Psi_L$;i.e., sum of leaf solute($\Psi_s$ ) and pressure( $\Psi_p$) potentials) across dehydration states (ideally, at 0.2-0.3 MPa intervals).

$$
\Psi_L = \Psi_s + \Psi_p
$$

First, the saturated water content(SWC) is estimated by extrapolating the linear relationship between leaf water content and leaf water potential before leaf turgor loss($\Psi_{tlp}$ ; i.e., point at which $\Psi_p=0$) to its y-intercept(0,SWC). Then, the water content relative to saturation( relative water content; RWC) is estimated for each leaf. From the linear relationship between 100-RWC and inverse leaf water potential, we estimate the osmotic or solute potential at full turgor as the y-intercept of that line and the pressure potential as $\Psi_L$-$\Psi_s$. Other parameters are estimated subsequently from the remaining portions of the curve below and above $\Psi_{tlp}$ . Many have attempted to create models from the curve (e.g., a logistic plus a linear curve) but this does not accurately explain the relationship between 100-RWC and leaf water potential .

The impetus for these models is the need to discern pressure volume curve parameters consistently across leaves, species and investigators. Common pressure volume curve protocols request that researchers estimate $\Psi_{tlp}$ by choosing the point of transition between the more curved (above turgor loss) and linear(below turgor loss) portions of the curve of 100-RWC and inverse leaf water potential. This can be assisted by choosing the point ranges below turgor loss, including at least five points, which have the greatest coefficient of variation. While this can be a very systematic method for producing PV parameters, selecting this point is crucial to the prediction of other parameters and has been found to vary among investigators testing the same data. Furthermore, the need to estimate these parameters for larger data sets may be time intensive reducing throughput.

Here we propose an improvement upon published protocols for pressure volume parameter estimation[@sack2011], which allows the automated estimation of the important inflection point around $\Psi_{tlp}$ . This script we have constructed has two primary assumptions: modulus of elasticity (\$\\epsilon\$) linearly declines above turgor loss point, and there are at least four points below turgor loss point. To estimate turgor loss point automatically, first we fit the linear relationship(using standard major axis) between inverse $\Psi_L$ and 100-RWC. The y-intercept here is the $\pi_O$ or \$\\Psi_s\$ when \$\\Psi_L=0\$. Multiplying $\pi_0$ by -1 is the maximum $\Psi_p$ because at full turgor $\Psi_L= 0$ or alternatively, $\Psi_s + \Psi_p =0$ . We fit $\Psi_L~(1-RWC)$ for the four driest points of the curve.

Next, we calculate $\Psi_p$ for each water potential:

```{r load data}

#establish environment
library(here)
library(tidyverse)
library(smatr)
library(readxl)

#functions
sma_slope<-function(x,#x variable often the water related variable
                    y){# y variable often leaf water potential 
  sd(x)/sd(y)
  
  }

sma_intercept<-function(x, y,#x and y variables x=water, y=water potential
                        slope){#single value estimated from the sma_slope function
  mean(x)-slope*mean(y)
}

rwd_tlp_solve<-function(intercept, slope){
  -(intercept/slope)
}
#load data
pv_dat<-read_excel(here("pvldcurvedata_10042022.xlsx"))%>%
  select(!c(fresh.weight.saturated, water.potential.bar, bag.weight,sp.leaf) )%>%
  mutate(inv.water.potential=-1/water.potential)

pp<-pv_dat%>%
  dplyr::filter(species=="quag_gj")#testing the method with quag

pp_leaf<-lapply(1:5, function(x) dplyr::filter(pp,leaf==x))#list of dfs, one for each leaf.

```

To estimate the saturated water content we use the intercept of the SMA relationship between leaf water and leaf water potential above and including tlp. The relationship is represented as: 

$$
Leaf~H_2O = \alpha * \Psi + b, 
$$
where $\alpha$ and b and the slope of the relationship estimated as: 

$$
\alpha=\frac{sd(leaf~water)}{sd(\Psi)}
$$
and, 

$$
b = mean(leaf~water)-\alpha*mean(\Psi)
$$

Relative water content(RWC) is then determined for each measurement as

$$
RWC=\frac{leaf~water}{SWC}
$$
```{r estimate SWC and RWC}

#select the four driest points for each leaf. 
pp_leaf_abovetlp<-lapply(1:5, function(x) pp_leaf[[x]]%>%
                            dplyr::arrange(desc(water.potential))%>%
                            slice_head(n=nrow(pp_leaf[[x]])-3))# doing three here because the line for SWC is often fit include the point right below TLP as well.

pp_leaf_swc<-data.frame(species=character(), leaf=numeric(), a_slope=numeric(), b_intercept=numeric())


for(leaf in 1:length(pp_leaf_abovetlp)){
  
  temp<-pp_leaf_abovetlp[[leaf]]  
  
  slope=sma_slope(x=temp$fresh.weight, y=temp$water.potential)
  intercept=sma_intercept(x=temp$fresh.weight, y=temp$water.potential, slope=slope)

  pp_leaf_swc[leaf,1]<-temp[1,1]
  pp_leaf_swc[leaf,2]<-temp[1,2]
  pp_leaf_swc[leaf,3]<-slope
  pp_leaf_swc[leaf,4]<-intercept

}

head(pp_leaf_swc)

#estimate relative water content 
pp_leaf_rwc<- lapply(1:5, function(leaf) pp_leaf[[leaf]]<-pp_leaf[[leaf]]%>%
                  mutate(saturated.water.content=pp_leaf_swc[leaf, "b_intercept"],
                         relative.water.content=fresh.weight/saturated.water.content,
                         relative.water.deficit=100-(relative.water.content*100)))


```

We then estimate the osmotic potential at full turgor using the slope of the SMA relationship between negative inverse of leaf water potential and relative water deficit below and including tlp. The relationship is represented as: 

$$
Leaf~H_2O = \alpha * \Psi + b, 
$$
where $\alpha$ and b and the slope of the relationship estimated as: 

$$
\alpha=\frac{sd(leaf~water)}{sd(\Psi)}
$$
and, 

$$
b = mean(leaf~water)-\alpha*mean(\Psi)
$$


```{r estimate line parameters below tlp}

##fit line to the points below TLP for 1/psiL vs 1-RWC

#select the four driest points for each leaf. 
pp_leaf_selectdry<-lapply(1:5, function(x) pp_leaf_rwc[[x]]%>%
                            dplyr::arrange(desc(water.potential))%>%
                            slice_tail(n=4))

#establish dataframe and estimate sma 
leaf_tlp_est<-data.frame(species=character(), leaf=numeric(), a_slope=numeric(), b_intercept=numeric())
  
for(leaf in 1:length(pp_leaf_selectdry)){
  
  temp<-pp_leaf_selectdry[[leaf]]  
 
    slope=-sma_slope(x=temp$ inv.water.potential, y=temp$relative.water.deficit)
  intercept=sma_intercept(x=temp$ inv.water.potential, y=temp$relative.water.deficit, slope=slope)

  leaf_tlp_est[leaf,1]<-temp[1,1]
  leaf_tlp_est[leaf,2]<-temp[1,2]
  leaf_tlp_est[leaf,3]<-slope
  leaf_tlp_est[leaf,4]<-intercept

}

pp_leaf_potentials<- lapply(1:5, function(leaf) pp_leaf_rwc[[leaf]]<-pp_leaf_rwc[[leaf]]%>%
                  mutate(osm.pot.fullturgor=-1/leaf_tlp_est[leaf, "b_intercept"],
                         max.pressure.potential=-osm.pot.fullturgor,
                         osmotic.potential=-1/(leaf_tlp_est[leaf, "b_intercept"]+leaf_tlp_est[leaf, "a_slope"]*relative.water.deficit),
                         pressure.potential=water.potential-osmotic.potential))
  

```

```{r estimate tlp}

##separate the above tlp points and estimate parameters for psi_p vs rwc
pres_abovetlp<-lapply(1:5, function(x) pp_leaf_potentials[[x]]%>%
                            dplyr::arrange(desc(water.potential))%>%
                            slice_head(n=nrow(pp_leaf_potentials[[x]])-3))

pres_param_abovetlp<-data.frame(species=character(), leaf=numeric(), a_slope=numeric(), b_intercept=numeric())
  
for(leaf in 1:length(pres_abovetlp)){
  
  temp<-pres_abovetlp[[leaf]]  
 
    slope=sma_slope(x=temp$pressure.potential, y=temp$relative.water.deficit)
    intercept=sma_intercept(x=temp$pressure.potential, y=temp$relative.water.deficit, slope=slope)

  pres_param_abovetlp[leaf,1]<-temp[1,1]
  pres_param_abovetlp[leaf,2]<-temp[1,2]
  pres_param_abovetlp[leaf,3]<-slope
  pres_param_abovetlp[leaf,4]<-intercept

}

##
pp_leaf_tlps<- lapply(1:5, function(leaf) pp_leaf_potentials[[leaf]]<-pp_leaf_potentials[[leaf]]%>%
                  mutate(relative.water.deficit.attlp=rwd_tlp_solve(pres_param_abovetlp[leaf, "b_intercept"], pres_param_abovetlp[leaf, "a_slope"]),
                         relative.water.content.attlp=100-relative.water.deficit.attlp,
                         leaf.water.potential.attlp=-1/(leaf_tlp_est[leaf, "a_slope"]*relative.water.deficit.attlp + leaf_tlp_est[leaf, "b_intercept"]),
                         modulus=max.pressure.potential/((100-relative.water.content.attlp)/100)))
  

```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
