---
title: "Package Test"
output: html_notebook
---

## Test the package functionality with data from ITV-PV

In this notebook, my aim is to validate the use of the functions within the [pvest](github.com/brownegm/pvest) package which takes raw measurements of leaf water potential and water mass of plant organs and estimates pressure volume curve parameters.

```{r load-packages, include=F}

# Create space
library(tidyverse)
library(here)
library(readxl)
library(pvest)

```

### Load data {#sec-data}

First, I load the raw values of the leaf water potential and leaf water mass. I only work with non-duplicates first for simplicity. Then, I create another object which loads the manually estimated parameters.

```{r load-data, echo=FALSE}
# load data

## individual mass and pv data
pv_dat<-readxl::read_xlsx(here("inst","extdata", "pvldcurvedata_10042022.xlsx"))%>%
  select(!c(fresh.weight.saturated, water.potential.bar, bag.weight,sp.leaf))%>%
  as.data.frame()

pvsps<-pv_dat%>%transmute(sp_indiv=paste0(toupper(species),leaf))# summarize and create list of species names 

#only the unique ids for the data 
unique.ids<-pvsps%>%
  dplyr::filter(pvsps$sp_indiv%in%itvpv$spcind)%>%
  pull(1)%>%
  unique()

## paper data summarized
itvpv <- readr::read_csv(here("inst/extdata/pvdat.csv"))

itvpv<-itvpv%>%
  filter(!grepl(c("PLRA|ENFA|BAGA|QUAG|RAIN|CEBE"), spcode))#filter these out; for the moment they add complexity

itvpv[itvpv$spcode=="CLLA","individual"]<-as.character(c(1:6))
itvpv[itvpv$spcode=="CLLA","spcind"]<-paste0("CLLA", c(1:6))

# match IDs of est output
itvpv<-itvpv%>%
  mutate(unique_id=paste(tolower(spcode),individual,sep = "_"), .before=swc)

itvpv_byspecies<-sumParams(itvpv, spcode)#summarize the manual estimates by species

head(pv_dat)
head(itvpv)

```

### Compute parameter estimates

Once the data is loaded estimate the parameters using the package function `estParams()` .

```{r compute-parameter-estimates}
# filter the unique ids that aren't in the summarized df
pv_dat_fil<-pv_dat%>%
  dplyr::filter(paste0(toupper(species),leaf) %in% unique.ids)

# compute
pv_params<-pvest::estParams(pv_dat_fil, fw.index = 5, wp.index = 4, dm.index = 3)

```
