---
title: "Estimation of Osmotic Variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimation of Osmotic Variables}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=F}
library(pvest)
library(dplyr)
```

The estimation of the osmotic variables (variables, units and symbols
are presented in Table 1) is the second step **after** estimating the
saturated water content and relative water content. These two
parameters, especially relative water content, are used to estimate the
osmotic variables using the linear relationship $\Psi_{leaf}=f(RWD)$.
See `RelativeWaterCD` for information on the estimation of those
parameters. The functions are an adaptation of the worksheet template
from [@Sack2011] and relationships drawn in [@Bartlett2012].

| Symbol         | Units    | Definition                                                                                                    |
|------------------------|------------------------|------------------------|
| $\pi_{o}$      | MPa      | Osmotic potential at full turgor.                                                                             |
| Max $\Psi_{p}$ | MPa      | Maximum pressure potential                                                                                    |
| $\Psi_s$       | MPa      | Osmotic potential for each hydration state                                                                    |
| $\Psi_{p}$     | MPa      | Pressure potential                                                                                            |
| $\Psi_{leaf}$  | MPa      | Leaf water potential. Sum of leaf osmotic and pressure potential                                              |
| a~f~           | \%       | Percent of total leaf water in the outside of cytoplasm                                                       |
| RWC~sym~       | g\*g^-1^ | Mass of water relative to full saturation within the symplast. Estimated with respect to apoplastic fraction. |

: **Table 1.** Variables, units and symbols of osmotic estimates.

## Estimate the $\pi_{o}$ and regression parameters with `OsmoticPotFullTurgor()`

`OsmoticPotFullTurgor()` estimates the osmotic potential at full turgor
from the linear relationship between relative water deficit and inverse
leaf water potential. This function itself is only used within the
`OsmoticEstimates()` function. This function estimates standard major
axis(SMA) regression parameters from the values **below** turgor loss
point. It assumes that your dataset is sufficient linear below turgor
loss point to approximate osmotic potential at full turgor and a
subsequent osmotic parameters. We assume that the data points at last
four hydration states all represent points below turgor loss point and
is the default.

The regression parameters are estimated by the helper
functions(`sma_slope` and `sma_intercept`). The relationship is
represented as:

$$
\Psi_{leaf}= \alpha * RWD + b, 
$$ where $\alpha$ and $b$ represent the slope and intercept of the
relationship and RWD is the relative water deficit:

$$
-\alpha=\frac{\sigma(RWD}{\sigma(\Psi_{leaf})}
$$ and,

$$
b = mean(RWD)-\alpha*mean(\Psi_{leaf})
$$

$$
RWD= 100-Relative~Water~Content
$$

### Usage

`OsmoticPotFullTurgor(data, fw.index, wp.index)`

-   **data** a data frame containing leaf water potential and leaf water
    mass

-   **wc.index** a numeric value indicating the column number where the
    relative water content data is within data frame

-   **wp.index** a numeric value indicating the column number where the
    leaf water potential(inverse) data is within data frame

### Output

The function outputs a list of three numbers: slope and intercept of the
line, and osmotic potential at full turgor.

## Estimate osmotic parameters using `OsmoticEstimates()`

The `OsmoticEstimates()` function estimates the osmotic potential at
full turgor ($\pi_{o}$) using `OsmoticPotFullTurgor()` and utilizes the
slope and intercept values to estimate the other parameters (Table 1).

The function first selects the subset of the last for values for a given
leaf then calculates the osmotic potential at full turgor and the
maximum pressure potential as :

$$
\pi_{o}= 1/intercept 
$$

$$
max~ \Psi_{p}= \pi_{o} * -1
$$

Because the leaf water potential at full hydration(\$\\Psi\_{leaf}=0\$)
is the sum of the osmotic potential at full turgor the maximum pressure
potential, you can solve for the maximum pressure potential by solving
for the $\Psi_{p}$ when $\Psi_{leaf}= 0$.

The osmotic potential at a given hydration state is estimated by
extrapolating the line of relative water deficit to inverse water
potential:

$$
\Psi_{s}=\frac{-1}{intercept+ (slope* RWD)}
$$

The pressure potential is the difference of the leaf water potential and
the leaf osmotic potential:

$$
\Psi_{p}=\Psi_{leaf}-\Psi_{s}
$$

Other parameters estimated are the apoplastic fraction and the
symplastic relative water content and deficit:

$$
a_{f}= 100 +  \frac{intercept}{slope}
$$

$$
RWC_{sym}=\frac{RWC-a_{f}}{100-a_{f}} *100
$$

$$
RWD_{sym}=100-RWC_{sym}
$$

### Usage

`OsmoticEstimate(data, wc.index, wp.index)`

-   **data** a data frame containing leaf water potential and leaf water
    mass

-   **wc.index** a numeric value indicating the column number where the
    relative water content data is within data frame

-   **wp.index** a numeric value indicating the column number where the
    inverse leaf water potential data is within data frame

## Example

```{r osm-estimation}
#load data for Quercus agrifolia leaf 1 
quag1<-pvest::data%>%
  filter(leaf==1)%>%
  mutate(inv.water.potential=-1/water.potential)

quag1_swc<- SaturatedWaterContent(data=quag1, 
                                  fw.index = 5, 
                                  wp.index = 4)

#create a data frame with the estimated SWC 
quag1_swc.df<-quag1

#create column that contains the saturated water content
quag1_swc.df$saturated.water.content<-rep(quag1_swc, nrow(quag1_swc.df))

#estimate RWC and RWD
quag1_rwc.est<-quag1_swc.df

quag1_rwc.est[,c("relative.water.content", "relative.water.deficit")]<-RelativeWaterCD(data=quag1_swc.df,
                           fw.index = 5)

q1_osm.est<-OsmoticEstimates(data=quag1_rwc.est, wc.index = "relative.water.deficit",wp.index = "inv.water.potential")

head(q1_osm.est%>%select(4:6, 10:16))
    
```

## References
